---
title: "ProjectOne"
output: html_document
---

#AMW: Nailed it.
# Data Project One

## 100 points
## DUE Sept 15 at 11:59 PM

Below are the questions for the first data practical assignment. This project uses the "FossilAnts.csv" file, located in the data directory for the project. The point value of each question is denoted next to it. A blank cell is below each for your answer; feel free to create more blank cells as needed.

0. (5 pts) Create a directory called `projects`, and in it, a subdirectory called `project_one`. Use `download.files` to get [this file](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/vignettes/project_one.Rmd), and [this one](https://raw.githubusercontent.com/BiologicalDataAnalysis2019/2023/main/projects/project_one/FossilAnts.csv) and save them to the `project_one` directory.


1. 5 pts. Import the `tidyverse` package and load the data. The data for this part of the practical is located in the data directory. Save the data in a variable called `project_dat`. Print the data to the screen to ensure it loaded correctly.

```{r}
project_dat <- read_csv("/cloud/project/projects/project_one/project_one_ants.csv")
```

2. (5pts) Check the datatypes of each column. There is a column called reference number. This is a static identifier - it should not be changed, and is an indexer used to identify this specimen uniquely. Do we want to treat it as an integer (this is an opinion question - answers may vary).

```
```{r}
I would not want it to treat it as an integer because the refrence number help us identify the specific data like a name. It should be treated as a string. 
```

```

3. (5 pts) Change the datatype of the reference number column to character. Take a peek at the function as.character().

```{r}
project_dat$reference_no <- as.character(project_dat$reference_no)
```

4. (5 pts) Look at your data. What are the missing data values? In particular, have a look at the `Tribe` column. In your opinion, are these intelligent missing values for the dataset? Why or why not? If not, how would you like to change them? 

```
```{r}
There is a mixture of "None" and "NAs" along with some actual intelligent values. The missing values should be reprsented by one thing bevause we dont know what the missing value means. 
```


```

6. (5 points) In the surveys dataset, we have genus and species split between two columns. Here, we have them combined. What are the pros and cons of the way we have recorded taxa in these two data sets.

```
```{r}
Combining Data:
Pro: 1. Reduced number of columns making it easier to work with.
Con: 1. Having them together means you cannot filter and  manupilate data easily.

Seperating the columns: 
Pro: 1. Flexibility to filter and work with different data. 
Con: 1. Adds columns and makes it slightly complicated. 
```
```


7. (5pts) Please look at the help page for the `separate` function.

```{r}
?separate
```


7. (10 pts) How could you separate one column into two? 

```{r}
project_dat %>% 

 separate(taxon,c("genus","species"))
```


```

Next, we will test a hypothesis. Your hypothesis is that there are more specimens in the 75 million years ago (mya) - 100 mya interval than the 30 mya to the present interval.

8. (5 pts) Write out the steps you would take to address this question. Will you need to split up the data? Will you need to group the data based on the values in some column? 

```{r}
1. Assign two variables, "data_75_100" for datas from 75 to 100 mya ago; the other variable named "data_35_present" for data from 35 million years ago till now. 
2. Count each data set for species. 
```


9. (15 pts) Perform the operations you described in (8). 

```{r}
data_35_present <- filter(project_dat,min_ma >= 0, max_ma <= 35)
#data_35_present displays 342 rows


data_75_100 <- filter(project_dat,min_ma >=75, max_ma<=100)
#data_75_100 dispays 48 rows
```

10. (5 pts) Do the results of your code support the hypothesis? 

```
It doesnt support the hypothesis because the data for 75 to 1000 mya shows way less rows than 35mya to present. 
```


11. (10 pts) Save the dataframe with the split taxon columns into a new directory called `project_one_data_output`. Save it as a csv file called "column_separated.csv" 

```{r}
 write.csv(project_dat, file = "/cloud/project/projects/project_one/project_one_data_output/column_separated.csv")
```

12. (10 pts) Produce a histogram of the number of specimens by minimum age. Look at the geom_histogram() help.

```{r}
ggplot(project_dat, aes(x= min_ma))+ geom_histogram()
```

13. (5 pts) Change the font size on your histogram so that it can be read comfortably on your computer screen from five feet away. Save the file as "large_font.pdf".

```{r}
large_font.pdf <- ggplot(project_dat, aes(x= min_ma))+ geom_histogram() + theme(text= element_text(size = 35))
ggsave("large_font.pdf", plot = large_font.pdf, width= 8, height = 6 )
```

14. Finally, produce a histogram of counts for each subfamily.

```{r}
project_subfamily <- project_dat %>% 
 group_by(subfamily) %>% 
 summarize(n=n())
 ggplot(project_subfamily, aes(x=n)) + geom_histogram()

```


# Graduate Students

1. You will do the above steps with the classroom dataset.
2. Next, you will choose three data steps from above to perform with your own data. This could be grouping, mutating, cleaning NA values, visualization, or any combination of the three. 
3. Why did you choose these steps? What function does this serve for your thesis and/or other research projects? 

